Magalhaes and Schwilk Supplementary Materials

### 2019-12-10 ms/Magalhaes_Schwilk_litter_flam_moisture.tex
joeR1

\noindent \textbf{Keywords:} California, flame spread rate, flammability, fuel consumption, fuel moisture, leaf litter, non-additivity

@@

With eight species, power to detect species-specific trait effects on these
moisture patterns was weak, but we examined one litter trait known to directly
influence flammability: litter bulk density and one leaf trait, specific leaf
area. Litter density is largely driven by leaf size in this system and, in
turn, controls the major axis of flammability variation in dry litter
\citep{Magalhaes+Schwilk-2012}. Our results suggest that moisture dynamics are
controlled by both traits intrinsic to the leaf particles (specific leaf area)
and an emergent property of the litter bed (bulk density). Specific leaf area
here is probably capturing some joint effect of cuticle thickness and total
surface area of litter particles, but we cannot test the morphological and
physical mechanisms of water absorption with these data. The oak has much
higher initial water retention than do the other species, followed by the
finest-needled pine, \emph{P. lambertiana}. Our study examines fresh litter,
but we expect that decomposition will affect moisture dynamics. For example,
Van Wagner \citep{Van_Wagner-1969}, investigating fuel weathering, found that
pine needles had a variation in moisture content 3 to 6 times greater when they
lacked their protective wax coating when compared to intact needles.

@@

\section*{Data and code accessibility}
All data and code for this project is at \url{https://github.com/schwilklab/trait-flam}

>>>>

\noindent \textbf{Keywords:} California, fire, flame spread rate, flammability, fuel consumption, fuel moisture, leaf litter, non-additivity

@@

With eight species, power to detect species-specific trait effects on these
moisture patterns was weak, but we examined one litter trait known to directly
influence flammability: litter bulk density and one leaf trait, specific leaf
area. Litter density is largely driven by leaf size in this system and, in
turn, controls the major axis of flammability variation in dry litter
\citep{Magalhaes+Schwilk-2012}. Our results suggest that moisture dynamics are
controlled by both traits intrinsic to the leaf particles (specific leaf area)
and an emergent property of the litter bed (bulk density). Specific leaf area
here is probably capturing some joint effect of cuticle thickness and total
surface area of litter particles, but we cannot test the morphological and
physical mechanisms of water absorption with these data. The oak has much
higher initial water retention than do the other species, followed by the
finest-needled pine, \emph{P. lambertiana}. Our study examines fresh litter,
but we expect that decomposition will affect moisture dynamics. For example,
Van Wagner \citeyear{Van_Wagner-1969}, investigating fuel weathering, found that
pine needles had a variation in moisture content 3 to 6 times greater when they
lacked their protective wax coating when compared to intact needles.

@@

\section*{Data and code accessibility}
All data and code for this project is at
\url{https://github.com/schwilklab/trait-flam}. DOI to be provided. For review,
use ``joe'' branch which contains this manuscript and code needed to create the
figures and tables.

### 2019-12-10 ms/SI.tex
joeR1

\begin{figure}[H]
  \centering
  \label{SI_fig2}
  \includegraphics[width=17cm]{figs_tables/SI_fig2_max_water_emmeans.pdf}
\caption{Estimated marginal means by species for maximum water retainability based on mixed model results. Shaded regions indicate 95\% confidence intervals for the marginal means.}
\end{figure}

\begin{figure}[H]
  \centering
  \label{SI_fig3}
\includegraphics[width=17cm]{figs_tables/SI_fig3_dessication_emmeans.pdf}
\caption{Estimated marginal means by species for dessication rate based on mixed model results. Shaded regions indicate 95\% confidence intervals for the marginal means.}
\end{figure}

\begin{figure}[H]
  \centering
\includegraphics[width=16cm]{figs_tables/SI_fig4_SLA_maxMC.pdf}
\caption{Maximum water retention by specific leaf area (SLA) across eight species.}
  \label{fig:maxmc-di}
\end{figure}

\begin{figure}[H]
  \centering
\includegraphics[width=16cm]{figs_tables/SI_fig5_di_bd.pdf}
\caption{Linear model predicting dessication rate as a function of specific leaf area and litter bulk density across eight species.}.
  \label{fig:bd-di}
\end{figure}

@@

% SI fig 5
\begin{figure}[H]
  \centering
\includegraphics[width=8cm]{figs_tables/fig2_SLA_maxMC.pdf}
\caption{Maximum water retention by specific leaf area across eight species.
  % ANOVA table in Table \ref{tab:mc_di_anova}.
}
  \label{fig:maxmc-di}
\end{figure}

% SI fig 6
\begin{figure}[H]
  \centering
\includegraphics[width=8cm]{figs_tables/fig3_di_bd.pdf}
\caption{Drying rate rate as a function of specific leaf area and litter bulk
  density across eight species.
  % ANOVA table in Table \ref{tab:mc_di_anova}
}.
  \label{fig:bd-di}
\end{figure}

>>>>

\begin{figure}[H]
  \centering
  \label{SI_fig2}
  \includegraphics[width=16cm]{figs_tables/SI_fig2_max_water_emmeans.pdf}
\caption{Estimated marginal means by species for maximum water retainability based on mixed model results. Shaded regions indicate 95\% confidence intervals for the marginal means.}
\end{figure}

\begin{figure}[H]
  \centering
  \label{SI_fig3}
\includegraphics[width=16cm]{figs_tables/SI_fig3_dessication_emmeans.pdf}
\caption{Estimated marginal means by species for dessication rate based on mixed model results. Shaded regions indicate 95\% confidence intervals for the marginal means.}
\end{figure}

\begin{figure}[H]
  \centering
\includegraphics[width=8cm]{figs_tables/SI_fig4_SLA_maxMC.pdf}
\caption{Maximum water retention by specific leaf area (SLA) across eight species.}
  \label{fig:maxmc-di}
\end{figure}

\begin{figure}[H]
  \centering
\includegraphics[width=8cm]{figs_tables/SI_fig5_di_bd.pdf}
\caption{Linear model predicting dessication rate as a function of specific leaf area and litter bulk density across eight species.}.
  \label{fig:bd-di}
\end{figure}

@@

% SI fig 5
\begin{figure}[H]
  \centering
\includegraphics[width=8cm]{figs_tables/SI_fig4_SLA_maxMC.pdf}
\caption{Maximum water retention by specific leaf area across eight species.
  % ANOVA table in Table \ref{tab:mc_di_anova}.
}
  \label{fig:maxmc-di}
\end{figure}

% SI fig 6
\begin{figure}[H]
  \centering
\includegraphics[width=8cm]{figs_tables/SI_fig5_di_bd.pdf}
\caption{Drying rate rate as a function of specific leaf area and litter bulk
  density across eight species.
  % ANOVA table in Table \ref{tab:mc_di_anova}
}.
  \label{fig:bd-di}
\end{figure}
